---
title: "Predictive Analytics - Regression"
author: "Kaia Huang"
date: "2026-02-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Regression Analysis to predict Boston House prices.

```{r}
# 1. Read the Boston House prices file.
housing <- read.csv("06_Regression_Boston_Housing_Prices.csv")
head(housing)
```

```{r}
summary(housing)
```

-Mean value (avarage) of Age of all the houses
```{r}
mean(housing$AGE, na.rm = TRUE)
```

- Data Cleaning (Replace missing values with 0's)
```{r}
library(tidyverse)
housing$AGE <- replace_na(housing$AGE, 0)
housing$CRIM <- replace_na(housing$CRIM, 0)
housing$ZN <- replace_na(housing$ZN, 0)
housing$INDUS <- replace_na(housing$INDUS, 0)
housing$CHAS <- replace_na(housing$CHAS, 0)
housing$LSTAT <- replace_na(housing$LSTAT, 0)
```

- Re-verify the mean value of age

```{r}
mean(housing$AGE, na.rm = TRUE)
```
- Multicollinearity: Correlation between Independent variables
- Correlation should be there between Independent variables and dependent variables

```{r}
library(ggcorrplot) 
# Correlation analysis between IVs and DVs 

correlation_matrix <- cor(housing) 

# ?ggcorrplot() 

ggcorrplot(correlation_matrix, 
           method = "square", 
           lab = T, 
           digits = 1, 
           type = "lower" ) 

```


```{r}
# 2. Training and Test split ---- 
# Sometimes validation sample is also used. 
#Use 70% data for training, 10% data for validation, 20% data for testing 

set.seed(1234) 
train_test_vector <- sample(c("Training", "Testing"), nrow(housing), replace = T, prob = c(0.7, 0.3)) 

# Create train dataset 

train <- housing[train_test_vector == "Training", ] 

# Create test dataset 

test <- housing[train_test_vector == "Testing", ] 
```


 - creat the model and summarize the model for coefficients
```{r}
# Select best model statistically by using the Adjusted R-Square 
lm_model <- lm(MEDV~., data=train) 
summary(lm_model) 
```
 
 
 - pick the best model using step-eise regression
```{r}
min_model <- lm(MEDV~1, data = train) 

max_model <- step(min_model,  
                  scope=(~ CRIM + ZN + INDUS + CHAS + NOX +  

                      RM + AGE + DIS + RAD + TAX + PTRATIO + B + LSTAT), 

                  direction = "forward", 

                  trace = 1) 

summary(max_model) 
```
 
 
 - capture Traning Error from the generated model
```{r}
#Setp 6: capture the training error (MSE)-. degree of freedom  
Training_Error <- sum((max_model$residuals)^2)/nrow(train)  
Training_Error 
```
 
 - Make predictions
```{r}
predictions <- predict(max_model,test[c(-14)]) # remove the actual outcome variable  # check data set and remove the column number e.g price of house   

test$Predicted_Output <- predictions  

testing_Error <-test$MEDV - test$Predicted_Output  

Testing_Error <- sum(testing_Error^2)/length(testing_Error)  

Testing_Error 

rmse_test <- sqrt(Testing_Error)
rmse_test
```
 
```{r}
library(caret) 
training_params <- trainControl(method = "repeatedcv",  

             number = 10, 

             repeats = 5, 

             verboseIter = T 

             ) 

# ?train() # multiple models, linear reg, logistic, SVM,  

model_lm1 <- train(MEDV ~.,  

                   data = train,  

                   method = "lm", 

                   trControl = training_params) 

summary(model_lm1) 
```
```{r}
library(car)
vif(max_model)
```
 
 
# There is no serial correlation between independent variables.
# vif – to check correlation.  
# The Variance Inflation Factor (VIF) measures how much the variance of an estimated regression coefficient is increased due to multicollinearity (correlation) among predictor variables 
# If VIF for a variable is 1, then it is not having any serial correlation with other IVs 
# 5 is acceptable 
# 10 should be absolutely avoided. 
 
```{r}
# Ridge Regression  

ridge_model <- train(MEDV ~.,  
               data =  train, 
               method = 'glmnet', 
               tuneGrid = expand.grid(alpha = 0,  
                                      lambda = seq(0, 5, length = 50)), # set alpha = 0 , lambda = 0 to 5 and set of 50  
                trControl = training_params) 
```

```{r}
ridge_model$bestTune 
```
 
```{r}
prediction <- predict(ridge_model,test[c(-14)])
head(predictions)
```
```{r}
testing_Error <-predictions - test$MEDV 
Testing_Error <- sum(testing_Error^2)/length(testing_Error)  
Testing_Error 

rmse_test <- sqrt(Testing_Error)
rmse_test
```
 
 