---
title: "Assignment - 07 - Simulating Roadside Assistance Requests"
author: "Kaia Huang"
date: "2026-04-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Justification for Using Poisson Distribution

Independence of Events: Each request for roadside assistance is an independent event. For example, getting a flat tire today does not change the probability of having a dead battery in the future.

Counting Occurrences: The Poisson distribution is ideal because we are counting the specific number of times a claim occurs within a fixed time interval (one year). In this scenario, we can only count the events that "happen," as it is impossible to count how many times an event did "not happen".

Randomness and Rarity: Roadside claims happen purely by chance. Given that we have a specific average rate of 0.5 claims per year, this fits the characteristics of a Poisson model perfectly for simulating business uncertainty.

```{r Q2. normal_distribution}
set.seed(123)  # Create a vector of 20 random numbers following a Poisson distribution with a mean (lambda) of 0.5. Name this vector simulated_claims. This represents the annual number of roadside assistance requests for a sample of 20 customers. 
num_samples <- 20
avg_claims <- 0.5         # lambda 
total_repair_cost <- 100  # Cost of repair through 3rd party 
agency_pay_ratio <- 0.5   # The insurance agency pays 50% cost to 3rd party per roadside assistance repair. 

# simulate 20 samples (for 0 to 19 claims) 
simulated_claims <- rpois(n = 20, lambda = 0.5) 

df_claims <- data.frame(claims = simulated_claims)

# Approach - 1
# Use base plot ()
sorted_claims <- sort(simulated_claims)
print(head(sorted_claims))

density_sample <- dpois(sorted_claims, lambda = 0.5)
plot(x = sorted_claims, 
     y = density_sample, 
     type = 'h', 
     main = "Poisson Distribution (lambda = 0.5)",
     xlab = "Number of Claims",
     ylab = "Probability")
points(x = sorted_claims, y = density_sample, pch = 16)


# Approach - 2
# Use geom_density (No need to calculate the densities separately)
library(ggplot2)
df_claims <- data.frame(claims = simulated_claims) +
   ggplot(df_claims, aes(x = claims)) +
  geom_bar(aes(y = ..prop..), fill = "skyblue", color = "black", width = 0.5) +
  labs(title = "Simulated Roadside Assistance Claims",
       x = "Number of Claims per Year",
       y = "Proportion of Customers") +
  theme_minimal()

```

```{r}
# Option 1: The insurance agency only charges a premium. 
# Assumptions:  1. Cost of repair through 3rd party is $100  2. Insurance premium is $50 per annum for roadside assistance.  3. The insurance agency pays 50% cost to 3rd party per roadside assistance repair.  4. Average number of claims for roadside assistance is 0.5 
premium1 <- 50 
profit_opt1 <- premium1 - (simulated_claims * total_repair_cost * agency_pay_ratio)
```

```{r}
# Option 2: The insurance agency charges a premium and a deductible.  
# Assumptions:  1. Cost of repair through 3rd party is $100  2. Insurance premium is $25 per annum for roadside assistance.  3. There is a deductible of $25 per roadside assistance repair.  4. The insurance agency pays 50% cost to 3rd party per roadside assistance repair.  5. Average number of claims for roadside assistance is 0.5 
premium2 <- 25
deductible2 <- 25
profit_opt2 <- premium2 + (simulated_claims * deductible2) - 
               (simulated_claims * total_repair_cost * agency_pay_ratio)
```

```{r}
# Option 3: The insurance agency charges a premium, a deductible, and limits the number of roadside assistance repairs that the customers can raise per annum.
# Assumptions:  1. Cost of repair through 3rd party is $100  2. Insurance premium is $25 per annum for roadside assistance.  3. There is a deductible of $25 per roadside assistance repair.  4.  The customer roadside assistance request will not be paid by the insurance agency after 2nd claim.  5. The insurance agency pays 50% cost to 3rd party per roadside assistance repair.  6. Average number of claims for roadside assistance is 0.5 

# The customer roadside assistance request will not be paid by the insurance agency after 2nd claim. 
premium3 <- 25 
deductible3 <- 25 
max_covered_claims <- pmin(simulated_claims, 2) 

# Profit
profit_opt3 <- premium3 + (simulated_claims * deductible3) - 
               (max_covered_claims * total_repair_cost * agency_pay_ratio)
```

```{r}
results <- data.frame(
  Sample_ID = 0:19, 
  Claims = simulated_claims,
  Profit_Opt1 = profit_opt1,
  Profit_Opt2 = profit_opt2,
  Profit_Opt3 = profit_opt3
)
```

```{r}
# Expected Profit 
expected_profits <- colMeans(results[, c("Profit_Opt1", "Profit_Opt2", "Profit_Opt3")])
print(expected_profits)
```

